Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-07 model.tex
fixing and adding forecast figures

Given a nonlinear model of a deterministic dynamical system in the
form of a delay-coordinate embedding like those in Figure~\ref{fig:embedding},
one can build deterministic forecast algorithms by capturing and
exploiting the geometry of the embedding.  Many techniques have been
developed by the dynamical systems community for this purpose
(e.g.,~\cite{weigend-book,casdagli-eubank92,Smith199250}).  Perhaps the most straightforward
is the ``Lorenz method of analogues'' (LMA), which is essentially
nearest-neighbor prediction in the embedded state
space~\cite{lorenz-analogues}.  Even this simple algorithm---which
builds predictions by finding the nearest neighbor in the embedded space of the given point, then taking that neighbor's path as the
forecast provides quite accurate forecasts when a process generates and transmits information in this way. In particular, when an observation in a time series has redundancy in the form of nonlinear structure this forecast scheme should be a great choice. However, since this method is based on near-neighbor prediction in a reconstructed state-space, if a system does not exhibit deterministic structure this method can do quite poorly. {\color{red}[[Joshua:Not a fan of this wording]]}
%% [[If space: Add a segue sentence: next section uses permutation
%% entropy to explore that conjecture.]]

@@

\begin{figure}[htbp]
  \centering
      \begin{subfigure}{0.48\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colMeanForecast}
    \caption{\col na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%
   \begin{subfigure}{0.48\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccMeanForecast}
    \caption{\gcc na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%      
  \\
    \begin{subfigure}{0.48\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colARIMAForecast}
    \caption{\col ARIMA}
    \label{fig:colARIMA}
  \end{subfigure}
  \begin{subfigure}{0.48\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccARIMAForecast}
    \caption{\gcc ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%
  \\

      \begin{subfigure}{0.48\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colLMAForecast}
    \caption{\col LMA}
    \label{fig:colLMA}
  \end{subfigure}
      \begin{subfigure}{0.48\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccLMAForecast}
    \caption{\gcc LMA}
    \label{fig:gccLMA}
  \end{subfigure}  
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   \caption{
{\color{red}[[Joshua: I should probably add \svdfive as it is a persistent example making this figure 9x9.}For each of these, we plot the predicted value $p_n$ against the correct value $c_n$. On this type of plot a perfect prediction lies exactly on the diagonal, that is the line $p_n = c_n$, e.g., \ref{fig:colLMA} is a near perfect prediction where-as \ref{fig:gccARIMA} is a very poor prediction. }\label{fig:gcc_vs_col}  
  %\end{subfigure}
\end{figure} 
In order to analyze correctness of each prediction we split each time series into two pieces: the first 90\% referred to as the ``learning" or ``training" signal, $\{X_{i,obs}\}_{i=1}^{n}$ and the last 10\% known as the ``test" or ``correct" signal $\{c_j\}_{j=n+1}^{k+n+1}$. The learning signal is used to train an initial model (e.g., LMA or ARIMA) as described in Section \ref{sec:model}. The test signal is used both to assess the models forecasting accuracy and for any refitting that may be necessary. In particular, we perform $k$ one-step predictions, after each one-step prediction\footnote{We would like to note that this rebuilding occurs due to a problem with ARIMA models converging to a mean prediction if too long of a prediction horizon is used, this is not a handicap of either LMA or na\"ive.} we append the training signal with the next point in the correct signal $c_j$, refit the model taking into account the new system measurement and perform another prediction. This is repeated $k$ times to obtain $\{p_j\}_{j=n+1}^{k+n+1}$.

@@

---works quite well on the trace in Figure~\ref{fig:ipc}, as
shown in Figure~\ref{fig:lmacol}. 
%
\begin{figure}
   \centering
\begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colPredShortTS}
    \caption{\col }
    \label{fig:lmacol}
  \end{subfigure}%

    \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colPredShortTS}
    \caption{\gcc}
    \label{fig:lmagcc}
  \end{subfigure} 
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colPredShortTS}
    \caption{\svdfive}
    \label{fig:lmasvdfive}
  \end{subfigure}% 
  \caption{{\color{red} [[If Liz thinks we should include these I actually need to generate this figure]]}An LMA-based forecast of
       processor-efficiency performance trace from the\col ,\gcc and \svdfive generating processes.  Red circles and blue $\times$s are the true and
       predicted values, respectively; vertical bars show where these
       values differ.}\label{fig:lmapredictions}
 \end{figure}
%
%\begin{figure}[htbp]
%  \centering
%   \includegraphics[width=\textwidth]{figs/colPredShortTS}
%    \caption{A forecast of the last 4,000 points of the %signal in
%      Figure~\ref{fig:ipc} using an LMA-based strategy on %the
%      embedding in Figure~\ref{fig:embedding}.  Red circles %and blue
%      $\times$s are the true and predicted values, %respectively;
%      vertical bars show where these values differ.}
%\label{fig:lmacol}
%\end{figure}
%
%\begin{figure}[htbp]
%  \centering
%    \includegraphics[width=\textwidth]{figs/colPredShortTS}
%     \caption{{\color{red} [[actually need to generate this %figure]]}An LMA-based forecast of the last 4,000 points of a
%       processor-load performance trace from the \gcc
%       benchmark.  Red circles and blue $\times$s are the %true and
%       predicted values, respectively; vertical bars show %where these
%       values differ.}
%\label{fig:lmagcc}
%\end{figure}
%
On the other hand, if we use the same approach to forecast the
processor efficiency (IPC) of the \gcc time series, the
prediction is far less accurate; see Figure~\ref{fig:lmagcc}.
Figure~\ref{fig:lmasvdfive} is clearly a better prediction than Figure~\ref{fig:gccLMA} but not nearly as good as the \col forecast.

>>>>

Given a nonlinear model of a deterministic dynamical system in the
form of a delay-coordinate embedding like those in Figure~\ref{fig:embedding},
one can build deterministic forecast algorithms by capturing and
exploiting the geometry of the embedding.  Many techniques have been
developed by the dynamical systems community for this purpose
(e.g.,~\cite{weigend-book,casdagli-eubank92,Smith199250}).  Perhaps the most straightforward
is the ``Lorenz method of analogues'' (LMA), which is essentially
nearest-neighbor prediction in the embedded state
space~\cite{lorenz-analogues}.  Even this simple algorithm---which
builds predictions by finding the nearest neighbor in the embedded space of the given point, then taking that neighbor's path as the
forecast, provides accurate forecasts when the generating process is a deterministic dynamical system. In particular, when an observation in a time series has redundancy in the form of nonlinear predictive structure this forecast scheme should be a great choice. However, since this method is based on near-neighbor prediction in a reconstructed state-space, if a system does not exhibit deterministic structure this method can do quite poorly. {\color{red}[[Joshua:Not a fan of this wording]]}
%% [[If space: Add a segue sentence: next section uses permutation
%% entropy to explore that conjecture.]]

@@

\begin{figure}[htbp]
  \centering
      \begin{subfigure}{0.33\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colMeanForecast.png}
    \caption{\col\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%
   \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccMeanForecast.png}
    \caption{\gcc\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}% 
     \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveMeanForecast.png}
    \caption{\svdfive\\ na\"ive }
    \label{fig:gccMEAN}
  \end{subfigure}%      
  \\
    \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colARIMAForecast.png}

    \caption{\col ARIMA}
    \label{fig:colARIMA}
  \end{subfigure}
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccARIMAForecast.png}
    \caption{\gcc \\ ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%
  \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveARIMAForecast.png}
    \caption{\svdfive \\ ARIMA }
    \label{fig:gccARIMA}
  \end{subfigure}%    
  \\

      \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/colLMAForecast.png}
    \caption{\col LMA}
    \label{fig:colLMA}
  \end{subfigure}
      \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/gccLMAForecast.png}
    \caption{\gcc LMA}
    \label{fig:gccLMA}
  \end{subfigure}  
        \begin{subfigure}{0.32\columnwidth}
    \includegraphics[width=\columnwidth]{figs/svdfiveLMAForecast.png}
    \caption{\svdfive LMA}
    \label{fig:gccLMA}
  \end{subfigure}
  %\begin{subfigure}{0.5\textwidth}
  %  \includegraphics[width=1.0\textwidth]{figs/LMA_vs_ARIMA}
   \caption{
{\color{red}[[Joshua: I should probably add \svdfive as it is a persistent example making this figure 9x9.}For each of these, we plot the predicted value $p_n$ against the correct value $c_n$. On this type of plot a perfect prediction lies exactly on the diagonal, that is the line $p_n = c_n$, e.g., \ref{fig:colLMA} is a near perfect prediction where-as \ref{fig:gccARIMA} is a very poor prediction. }\label{fig:gcc_vs_col}  
  %\end{subfigure}
\end{figure} 
In order to analyze correctness of each prediction we split each time series into two pieces: the first 90\% referred to as the ``learning" or ``training" signal, $\{X_{i,obs}\}_{i=1}^{n}$ and the last 10\% known as the ``test" or ``correct" signal $\{c_j\}_{j=n+1}^{k+n+1}$. The learning signal is used to train an initial model (e.g., LMA or ARIMA) as described in Section \ref{sec:model}. The test signal is used both to assess the models forecasting accuracy and for any refitting that may be necessary. In particular, we perform $k$ one-step predictions, after each one-step prediction\footnote{We would like to note that this rebuilding occurs due to a problem with ARIMA models converging to a mean prediction if too long of a prediction horizon is used, this is not a handicap of either LMA or na\"ive.} we append the training signal with the next point in the correct signal $c_j$, refit the model taking into account the new system measurement and perform another prediction. This is repeated $k$ times to obtain $\{p_j\}_{j=n+1}^{k+n+1}$.

@@

---works quite well on the trace in Figure~\ref{fig:ipc}, as
shown in Figure~\ref{fig:lmacol}. 
%
%\begin{figure}
%   \centering
%\begin{subfigure}{\columnwidth}
%    \includegraphics[width=\columnwidth]{figs/%colPredShortTS}
%    \caption{\col }
%    \label{fig:lmacol}
%  \end{subfigure}%  
%  
%    \begin{subfigure}{\columnwidth}
%    \includegraphics[width=\columnwidth]{figs/%colPredShortTS}
%    \caption{\gcc}
%    \label{fig:lmagcc}
%  \end{subfigure} 
%  \begin{subfigure}{\columnwidth}
%%    \includegraphics[width=\columnwidth]{figs/%colPredShortTS}
%    \caption{\svdfive}
%    \label{fig:lmasvdfive}
%  \end{subfigure}% 
%  \caption{{\color{red} [[If Liz thinks we should %include these I actually need to generate this %figure]]}An LMA-based forecast of
%       processor-efficiency performance trace from the%\col ,\gcc and \svdfive generating processes.  Red %circles and blue $\times$s are the true and
%       predicted values, respectively; vertical bars %show where these
%       values differ.}\label{fig:lmapredictions}
% \end{figure}
%
%\begin{figure}[htbp]
%  \centering
%   \includegraphics[width=\textwidth]{figs/colPredShortTS}
%    \caption{A forecast of the last 4,000 points of the %signal in
%      Figure~\ref{fig:ipc} using an LMA-based strategy on %the
%      embedding in Figure~\ref{fig:embedding}.  Red circles %and blue
%      $\times$s are the true and predicted values, %respectively;
%      vertical bars show where these values differ.}
%\label{fig:lmacol}
%\end{figure}
%
%\begin{figure}[htbp]
%  \centering
%    \includegraphics[width=\textwidth]{figs/colPredShortTS}
%     \caption{{\color{red} [[actually need to generate this %figure]]}An LMA-based forecast of the last 4,000 points of a
%       processor-load performance trace from the \gcc
%       benchmark.  Red circles and blue $\times$s are the %true and
%       predicted values, respectively; vertical bars show %where these
%       values differ.}
%\label{fig:lmagcc}
%\end{figure}
%
On the other hand, if we use the same approach to forecast the
processor efficiency (IPC) of the \gcc time series, the
prediction is far less accurate; see Figure~\ref{fig:lmagcc}.
Figure~\ref{fig:lmasvdfive} is clearly a better prediction than Figure~\ref{fig:gccLMA} but not nearly as good as the \col forecast.
